Active-contours driven registration applied in the correction of susceptibility-derived distortions of diffusion MR images

### 2015-04-05 Publications/2015-NeuroImage/03-introduction.tex
my fixes

Early attempts to delineate the \gls*{wm} proceeded by thresholding the 
  \gls*{fa}\footnote{the \gls*{fa} is a scalar parameter of diffusion derived from
  the \gls*{dmri} data.} map.
However, the mask and subsequent analyses strongly depend on the threshold set
  \citep{taoka_fractional_2009} owing to the large variability of \gls*{fa} and its
  decay with the increase of incoherence between fibers.
\cite{zhukov_level_2003} proposed active contours with edges represented
  by level sets, evolving on directionally invariant scalar maps.
\cite{rousson_level_2004} successfully segmented the corpus callosum with
  region-based level-sets on the eigenvalues of tensors fitted on the
  \gls*{dmri} data.
A large body of work targets the definition of appropriate features to segmentation,
  such as the 5D representation proposed by \cite{jonasson_segmentation_2005}.
Other efforts include mixture models on sets of directionally invariant maps
  \citep{liu_brain_2007}, iterative \citep{hadjiprocopis_unbiased_2005} and
  hierarchical \citep{lu_segmentation_2008} clusterings,
  and graph-cuts \citep{han_experimental_2009},
  and volume fraction modeling \citep{kumazawa_improvement_2013}.

@@

Therefore, given the described circumstances, a method performing simultaneously
  segmentation in \gls*{dmri} space and registration of the corresponding \gls*{t1} image
  to \gls*{dmri} data could provide the optimal solution.
Previously, joint segmentation and registration has been successfully applied to
  longitudinal object tracking \citep{paragios_level_2003} and atlas-based
  segmentation \citep{gorthi_active_2011}.
The most extended approach is integrating active contours in image registration 
  methods.
\cite{unal_coupled_2005} and later \cite{wang_joint_2006},
  improved an existing method \citep{yezzi_variational_2003} based on linear registration
  to the nonlinear case implementing a free-form deformation field.
\cite{droske_mumfordshah_2009} reviewed the latter set of techniques, and proposed two different
  approaches to apply the Mumford-Shah functional \citep{mumford_optimal_1989} in simultaneous
  registration and segmentation, through the propagation of the deformation field from
  the contours to the whole image definition.
Recently, \cite{guyader_combined_2011} proposed a 2D simultaneous segmentation and
  registration method using level sets and a nonlinear elasticity smoother on the
  displacement vector field, which preserves topology even with very large deformations.
Finally, \cite{gorthi_active_2011} extended the existing methodologies using a multiphase
  level-set function for the registration of several active contours, in the application
  of atlas-based segmentation.

The hypothesis under this research is that the registration and segmentation
  problems in \gls*{dmri} can be solved simultaneously, increasing the geometrical
  accuracy of the process.
We propose a tool called \emph{regseg} to exploit the prior information of shapes
  extracted on \gls*{t1} images to register the anatomical reference
  to \gls*{dmri} space, implicitly segmenting these data.
The approach uses active contours without edges \citep{chan_active_2001} that evolve driving a
  free-deformation field of B-Spline functions.
Optimization is performed using a descent strategy of explicit shape-gradients
  \citep{besson_dream2s_2003,herbulot_segmentation_2006}, therefore \emph{regseg}
  does not implement levelsets as most of the presented methods.
The nonlinear warping caused by distortion is aligned with one of the imaging axis due
  to the theory underpinning \gls*{epi} sequences.
Therefore, \emph{regseg} includes an anisotropic regularizer of the displacement field 
  proposed by \cite{nagel_investigation_1986}.
Finally, we present an instrumentation framework to evaluate correction and
  segmentation pipelines using realistic distortions as ground truth.
With the integration of \emph{regseg} and an in-house implementation of the 
  \gls*{t2b} correction in the framework, we compare the performance
  of both solutions.

>>>>

Early attempts to delineate the \gls*{wm} proceeded by thresholding the 
  \gls*{fa}\footnote{the \gls*{fa} is a scalar parameter of diffusion derived from
  the \gls*{dmri} data.} map.
However, the mask and subsequent analyses substantially depend on the threshold set
  \citep{taoka_fractional_2009} owing to the large variability of \gls*{fa} and its
  decay with the increase of incoherence between fibers.
\cite{zhukov_level_2003} proposed active contours with edges represented
  by levelsets, evolving on directionally invariant scalar maps.
\cite{rousson_level_2004} successfully segmented the corpus callosum with
  region-based levelsets on the eigenvalues of tensors fitted on the
  \gls*{dmri} data.
A large body of work targeted the definition of appropriate features to cluster,
  such as the 5D representation proposed by \cite{jonasson_segmentation_2005}.
Other efforts include mixture models on sets of directionally invariant maps
  \citep{liu_brain_2007}, iterative \citep{hadjiprocopis_unbiased_2005} and
  hierarchical \citep{lu_segmentation_2008} clusterings,
  and graph-cuts \citep{han_experimental_2009},
  and volume fraction modeling \citep{kumazawa_improvement_2013}.

@@

Therefore, given the described circumstances, a method performing simultaneously
  segmentation in \gls*{dmri} space and registration of the corresponding \gls*{t1} image
  to the \gls*{dmri} data could provide the optimal solution.
Previously, joint segmentation and registration has been successfully applied to
  longitudinal object tracking \citep{paragios_level_2003} and atlas-based
  segmentation \citep{gorthi_active_2011}.
The most extended approach is integrating active contours in image registration 
  methods.
\cite{unal_coupled_2005} and later \cite{wang_joint_2006},
  improved an existing method \citep{yezzi_variational_2003} based on linear registration
  to the nonlinear case implementing a free-form deformation field.
\cite{droske_mumfordshah_2009} reviewed the existing techniques, and proposed two different
  approaches to apply the Mumford-Shah functional \citep{mumford_optimal_1989} in simultaneous
  registration and segmentation, through the propagation of the deformation field from
  the contours to the whole image definition.
Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
  registration method in 2D using levelsets and a nonlinear elasticity smoother on the
  displacement vector field, which preserves topology even with very large deformations.
Finally, \cite{gorthi_active_2011} extended the existing methodologies using a multiphase
  levelset function for the registration of several active contours, in the application
  of atlas-based segmentation.

The hypothesis under this research is that the registration and segmentation
  problems in \gls*{dmri} can be solved simultaneously, increasing the geometrical
  accuracy of the process.
We propose a tool called \emph{regseg} to exploit the prior information of shapes
  extracted on \gls*{t1} images to register the anatomical reference
  to \gls*{dmri} space, implicitly segmenting these data.
The approach uses active contours without edges \citep{chan_active_2001} that evolve driving a
  free-form deformation field of B-Spline basis functions.
Optimization is performed using a descent strategy of explicit shape-gradients
  \citep{besson_dream2s_2003,herbulot_segmentation_2006}, therefore \emph{regseg}
  does not implement levelsets as most of the presented methods.
Since the nonlinear distortion is aligned with one of the imaging axis (see 
  \autoref{sec:datasets}), \emph{regseg} includes an anisotropic regularizer of
  the displacements field proposed by \cite{nagel_investigation_1986}.
Finally, we evaluate \emph{regseg} using an extension of our instrumentation framework
  \citep{esteban_simulationbased_2014} that simulates known and realistic distortions
  on \gls*{dmri} data.
The evaluation includes a comparison to the \gls*{t2b} correction by integrating in the framework
  an in-house implementation of the method.

### 2015-04-05 Publications/2015-NeuroImage/04-methods.tex
my fixes

\subsection{Cost-function derivation}\label{sec:methods_map}
%
In a Bayesian framework for registration \citep{wyatt_map_2003,pohl_bayesian_2006,gass_simultaneous_2014},
  the mappings $U$ in \eqref{eq:transform} are
  evaluated based on their posterior probability given the observed data
  $M$.
Let $\omegaset = \{\Omega_l: l \in \mathbb{N}, l \leq L\}$ be the set of $L$ competing regions in
  which $M$ is partitioned by the projection of \gammaset, 
  the given surfaces serving as shape priors.
Using the Bayes' rule, the posterior likelihood is computed as:

@@

  \begin{equation}
  P(M \mid U) = \underset{l}{\prod} \underset{\vec{r}\in \Omega_l}{\prod}
    P\left( \vec{f}' \mid U \right),
  \label{eq:bayes_aposteriori}
  \end{equation}
%
  where $\vec{f}' = M(\vec{r}')$ is the feature vector at the displaced
  position $\vec{r}'$ \eqref{eq:transform} in the multispectral target
  volume.
For convenience, and because it has been shown to be an appropriate approximation
  \citep{leemput_automated_1999,cuadra_comparison_2005}, we introduce two assumptions for each
  region $\Omega_l$:
  1) the features are i.i.d.; and 2) they can be modeled by multivariate normal
  distributions with parameters $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
  for each region $\Omega_l$.
We then insert the normal distribution into \eqref{eq:bayes_rule} to obtain the full
  formulation of the \emph{data term} of our registration model:

 	\begin{align}
  P( M \mid U) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) \notag\\
  &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
  \mdist{f'}{l} \right)}},
  \label{eq:pdf}
  \end{align}
%
  using $\mdist{f}{l}$ to denote the squared \emph{Mahalanobis distance} of $\vec{f}$ with respect
  to the descriptors of region $l$ as
  $\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.
$C$ is the number of channels comprised in the reference image $M$.

@@

  \begin{align}
  P(U) = \underset{\vec{r}}{\prod}\, p(\vec{u}) &=
  \underset{\vec{r}}{\prod}\, p_0(\vec{u}) \, p_1(\vec{u}), \\
  p_0(\vec{u}) &= \mathcal{N}( \vec{u} \mid 0, \mathbf{A}^{-1}), \notag\\
  p_1(\vec{u}) &= \mathcal{N}(  \nabla \vec{u} \mid 0, \mathbf{B}^{-1}),
  \label{eq:thikonov}
  \end{align}
%
  imposing that the distortion and its gradient have zero
  mean and variance governed by the matrices $\mathbf{A}$ and $\mathbf{B}$.
Finally, the maximum a posteriori problem is adapted into a variational one where we look for
  the minimum of an energy functional, by applying a log-transform:

  \begin{align}
  E(M \mid U) &= -\log \underset{l}{\prod}
  \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} \left( \vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \right)\,p_0( \vec{u})\,p_1( \vec{u}) \notag\\
  &= \const + \underset{l}{\sum} \int_{\Omega_l}
  \mdist{f'}{l} \,d\vec{r} \, + \notag\\
  &+ \int_{\Omega} \frac12 \left[ \vec{u}^T \mathbf{A} \vec{u} + (\nabla \vec{u})^T \mathbf{B} (\nabla \vec{u}) \right] \,d\vec{r}.
  \label{eq:energy}
  \end{align}
%

This expression is dual of the Mumford-Shah functional corresponding
  to a discrete \acrlong*{acwe} framework \citep{chan_active_2001}
  with the corresponding anisotropic regularization term of \cite{nagel_investigation_1986}.

@@

\paragraph*{Deformation model}\label{sec:deformation_model}
Since the vertices of the surfaces $\{\vec{v}_i: \vec{v}_i \subset \gammaset \}_{i=1 \ldots N_v}$
  likely lay off-grid, it is required to derive $u(\vec{v}_i)$ from a discrete set of parameters
  $\{\vec{u}_k\}_{k=1 \ldots K}$.
Densification is achieved through a set of associated basis functions $\psi_k$.
In our implementation, $\psi_k$ is chosen to be a tensor-product B-Spline kernel
  of degree 3: % ($\beta_3$).

  \begin{equation}
  \vec{v}_i' = \vec{v}_i + u(\vec{v}_i) = \vec{v}_i + \sum_k \psi_k(\vec{r}) \: \vec{u}_k.
  \label{eq:nodes_tfm}
  \end{equation}

@@

  \begin{equation}
  \frac{\partial u(\vec{r},t)}{\partial t} \propto - \frac{\partial E(M \mid U)}{\partial \vec{u}_k},
  \label{eq:general_gradient_descent}
  \end{equation}
%
  with $t$ being an artificial time parameter of the contour
  evolution, and $\vec{u}_k$ the parameters supporting the estimate
  $\hat{U}$ of the transformation at the current time point.
Let us assume that anisotropy is aligned with the imaging axes to simplify
  \eqref{eq:energy} into expression \eqref{eq:app_energy} of \ref{app:reg_term}
  and compute the derivative \eqref{eq:general_gradient_descent}:

@@

\subsection{Experiments and evaluation}
\label{sec:experiments_evaluation}
%
In order to demonstrate the performance of \emph{regseg}, we first conducted a battery of
  accuracy tests over synthetic data, and then applied to real \gls*{dmri} data.
\autoref{fig:evworkflows} presents the workflow implementing the evaluation instruments.
Besides a visual assessment of the results, we report quantitative evaluations using
  two metrics.
In the case of the phantoms, since we had produced distortions along the three
  available dimensions, we computed the Haussdorf distance, by reusing the
  ``point-to-cell'' method of \cite{commandeur_vtk_2011}.
Conversely, the susceptibility-derived distortions only happen along the \gls*{pe}
  axis of the image.
Therefore, a \gls*{swindex} can be computed as the one-to-one distance between corresponding
  vertices of surfaces, weighted by their respective Voronoi area $a_i$:

  \begin{equation}
  sWI = \frac{1}{\sum_i a_i} \sum\limits_i^{N_p} a_i\,\|
  \vec{v}_i - \hat{\vec{v}}_i \|
  \label{eq:swindex}
  \end{equation}
%
  where $\vec{v}_i$ are the locations of the $N_p$ vertices in each $p \in \{1, \dots P\}$
  prior, and $\hat{\vec{v}}_i$ is the location
  recovered corresponding to the vertex $\vec{v}_i$.

\subsection{Image Data and preprocessing}
\label{sec:datasets}

\paragraph*{Simulated phantoms}%
\label{sec:digital_phantoms}
We proved the concept on simplistic digital phantoms which we will name after their
  appearance as ``box'', ``ball'', ``L-shape'', and ``gyrus'' (\autoref{fig:phantom},
  boxes A and B).
Using \emph{phantomas} \citep{caruyer_phantomas_2014}, we simulated
  \gls*{t1} (TE/TR=10/1500ms) and \gls*{t2} images (TE/TR=90/5000ms)
  corresponding to each phantom type, with two resolutions each
  (1.0mm and 2.0mm isotropic).
The field of view at both resolutions was \isores[mm]{100}.
Simulations were corrupted with rician noise for a \gls*{snr} of 300.0.
The reference surfaces were extracted at high resolution, using 
  \texttt{mri\_tessellate}\footnote{A marching-cubes algorithm shipped with 
  \emph{FreeSurfer}.}.

\begin{figure*}
\includegraphics[width=\linewidth]{figures/figure03}
\caption{Evaluation workflow of phantoms}\label{fig:evphantoms}
\end{figure*}

To warp the simulated datasets, we generated 150 different realizations of
  the distortion $U_{true}$.
Smoothness is introduced using two levels of B-Spline functions, with control points evenly
  located in isotropic grids covering the full extent of the phantom.
Separation between control points was isotropic, and set to 50.5mm for the first level
  of distortion and 25.25mm for the second.
The rationale behind this choice is producing large deformations on the contours with
  the first level, and matching the properties of susceptibility-derived distortions
  as previously reported by \cite{irfanoglu_susceptibility_2011} with the finer grid.
The coefficients of the B-Spline functions were randomly generated for each resolution level.
Invertibility of $U_{true}$ is ensured by controlling the maximum displacement at each
  level (20.2mm and 10.1mm respectively) as studied in \citep{rueckert_diffeomorphic_2006}.

\paragraph*{Real datasets} %
\label{sec:human_connectome}
%
For the evaluation of the algorithm on real \gls*{dmri} data of human brains,
  we collected 16 datasets from the ``minimally preprocessed''
	 database of the \gls*{hcp}.
We refer the reader to \citep{essen_human_2012} for exact details about acquisition
  parameters, and \citep{glasser_minimal_2013} for the preprocessing issues.
The datasets comprehend a large set of images, containing \gls*{t1}, \gls*{t2} and
  multi-shell \gls*{dmri} images.
The original acquisitions are released within ``unprocessed'' packages, and
  the ``minimally preprocessed'' are corrected for artifacts, brain-extracted
  and spatially normalized, along with some results of the standard processing
  pipeline of \emph{FreeSurfer}.

@@

>>>>

\subsection{Cost-function derivation}\label{sec:methods_map}
%
In a Bayesian framework for registration \citep{wyatt_map_2003,pohl_bayesian_2006,gass_simultaneous_2014},
  the mappings $U$ in \eqref{eq:transform} are
  evaluated based on their posterior probability given the observed data
  $M$.
Let $\omegaset = \{\Omega_l: l \in \mathbb{N}, l \leq N_L\}$ be the set of $N_L$ competing regions in
  which $M$ is partitioned by the projection of \gammaset.
Using the Bayes' rule, the posterior likelihood is computed as:

@@

  \begin{equation}
  P(M \mid U) = \underset{l}{\prod} \underset{\vec{r}\in \Omega_l}{\prod}
    P\left( \vec{f}' \mid U \right),
  \label{eq:bayes_aposteriori}
  \end{equation}
%
  where $\vec{f}' = M(\vec{r}')$ is the feature vector at the displaced
  position $\vec{r}'$ \eqref{eq:transform} in the multispectral target
  volume.
For convenience, and because it has been shown to be an appropriate approximation
  \citep{leemput_automated_1999,cuadra_comparison_2005}, we introduce two assumptions for each
  region $\Omega_l$:
  1) the features are i.i.d.; and 2) they can be modeled by multivariate normal
  distributions with parameters $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
  for each region $\Omega_l$:

 	\begin{align}
  P( M \mid U) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) = \notag\\
  &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
  \mdist{f'}{l} \right)}},
  \label{eq:pdf}
  \end{align}
%
  using $\mdist{f}{l}$ to denote the squared \emph{Mahalanobis distance} of $\vec{f}$ with respect
  to the descriptors of region $l$ as
  $\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.
$C$ is the number of channels comprised in the reference image $M$.

@@

  \begin{align}
  P(U) = \underset{\vec{r}}{\prod}\, p(\vec{u}) &=
  \underset{\vec{r}}{\prod}\, p_0(\vec{u}) \, p_1(\vec{u}), \\
  p_0(\vec{u}) &= \mathcal{N}( \vec{u} \mid 0, \mathbf{A}^{-1}), \notag\\
  p_1(\vec{u}) &= \mathcal{N}(  \nabla \vec{u} \mid 0, \mathbf{B}^{-1}),
  \label{eq:thikonov}
  \end{align}
%
  imposing that the distortion and its gradient have zero
  mean and variance governed by the matrices $\mathbf{A}$ and $\mathbf{B}$.
Finally, the maximum a posteriori problem is adapted into a variational one where we look for
  the minimum of an energy functional, by applying $E(\vec{u}) = -\log P(U)$:

  \begin{align}
  E(\vec{u}) &= -\log \underset{l}{\prod}
  \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} \left( \vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \right)\,p_0( \vec{u})\,p_1( \vec{u}) \notag\\
  &= \const + \underset{l}{\sum} \int_{\Omega_l}
  \mdist{f'}{l} \,d\vec{r} \, + \notag\\
  &+ \int_{\Omega} \frac12 \left[ \vec{u}^T \mathbf{A} \vec{u} + (\nabla \vec{u})^T \mathbf{B} (\nabla \vec{u}) \right] \,d\vec{r}.
  \label{eq:energy}
  \end{align}
%
This expression is dual of the Mumford-Shah functional corresponding
  to a framework of \acrlong*{acwe} \citep{chan_active_2001}
  with the corresponding anisotropic regularization term of \cite{nagel_investigation_1986}.

@@

\paragraph*{Deformation model}\label{sec:deformation_model}
Since the vertices of the surfaces $\{\vec{v}_i: \vec{v}_i \subset \gammaset \}_{i=1 \ldots N_v}$
  likely lay off-grid, it is required to derive $\vec{u}_i = \vec{u}(\vec{v}_i)$ from a discrete set of parameters
  $\{\vec{u}_k\}_{k=1 \ldots K}$.
Densification is achieved through a set of associated basis functions $\psi_k$ \eqref{eq:nodes_tfm}.
In our implementation, $\psi_k$ is chosen to be a tensor-product B-Spline kernel
  of degree 3.

  \begin{equation}
  \vec{v}_i' = \vec{v}_i + \vec{u}_i = \vec{v}_i + \sum_k \psi_k(\vec{r}) \: \vec{u}_k.
  \label{eq:nodes_tfm}
  \end{equation}

@@

  \begin{equation}
  \frac{\partial \vec{u}(\vec{r},t)}{\partial t} \propto - \frac{\partial E(\vec{u})}{\partial \vec{u}_k},
  \label{eq:general_gradient_descent}
  \end{equation}
%
  with $t$ being an artificial time parameter of the contour
  evolution, and $\vec{u}_k$ the parameters supporting the estimate
  $\hat{U}$ of the transformation at the current time point.
Let us assume that anisotropy is aligned with the imaging axes to simplify
  \eqref{eq:energy} into expression \eqref{eq:app_energy} of \ref{app:reg_term}
  and compute its derivative \eqref{eq:general_gradient_descent}:

@@

\subsection{Data and corresponding evaluations}
\label{sec:datasets}
In order to assess the performance of \emph{regseg}, we defined a general
  evaluation protocol:
1) Extraction of the set of undistorted surfaces \gammaset{};  
2) computation of a ground-truth field of displacements that is applied to
  generate warped images ($M$) to segment;
3) execution of \emph{regseg} with \gammaset{} and the warped data as inputs; and
4) visual assessment and computation of error metrics.

\paragraph*{Simulated phantoms}\label{sec:digital_phantoms}
The workflow to generate the digital phantoms and assess
  \emph{regseg} on them is presented in \autoref{fig:evphantoms}.
A set of four binary objects (called ``box'', ``ball'', ``L'',
  and ``gyrus'') are generated combining binarization of
  analytical shapes and mathematical morphology.
Using \emph{FreeSurfer} tools, the reference surfaces \gammaset{}
  are generated.
The ground-truth distortion is generated with a chain of two displacement 
  fields supported by grids of B-Spline basis functions.
The coefficients of the basis functions are generated randomly for
  both levels in their three dimensions.
The three components of the displacements $\vec{u} = (u_d)$ 
  are bounded above by 40\% of the separation between control points
  at each level, to obtain diffeomorphic transforms
  after concatenation \citep{rueckert_diffeomorphic_2006}.
The first deformation field is intended to generate large warpings,
  with control points separated by 50.50mm along the three dimensions
  ($u_d\leq$ 20.20mm).
With the second warping, we look for a resulting field with smoothness
  close to the found in a typical distortion field of \gls*{dmri} data
  \citep{irfanoglu_susceptibility_2011}.
Therefore, the control points are separated 25.25mm ($u_d\leq$ 10.10mm).
Once the ground-truth deformation is generated, the original surfaces
  are warped interpolating the displacements field at each vertex.
The warped surfaces are binarized to generate a distorted version of
  the model in two target resolutions: \isores[mm]{2.0} and \isores[mm]{1.0}.
Using a \gls*{mr} simulator \citep{caruyer_phantomas_2014}, we synthesized
  \gls*{t1} (TE/TR=10/1500ms) and \gls*{t2} images (TE/TR=90/5000ms)
  corresponding to each phantom type, with two resolutions each
  (1.0mm and 2.0mm isotropic).
The field of view at both resolutions was \isores[mm]{100}.
Simulations were corrupted with rician noise for a \gls*{snr} of 300.0.

To evaluate the error, we computed the Haussdorf distance (as implemented
 by \cite{commandeur_vtk_2011}) between the surfaces warped with the ground-truth
 and their corresponding contours estimated with \emph{regseg}.

\begin{figure*}
\includegraphics[width=\linewidth]{figures/figure03}
\caption{Evaluation workflow of phantoms.

}\label{fig:evphantoms}
\end{figure*}

\paragraph*{Real datasets} %
\label{sec:human_connectome}
% 
\autoref{fig:evworkflows}
For the evaluation of the algorithm on real \gls*{dmri} data of human brains,
  we collected 16 datasets from the ``minimally preprocessed''
	 database of the \gls*{hcp}.
We refer the reader to \citep{essen_human_2012} for exact details about acquisition
  parameters, and \citep{glasser_minimal_2013} for the preprocessing issues.
The datasets comprehend a large set of images, containing \gls*{t1}, \gls*{t2} and
  multi-shell \gls*{dmri} images.
The original acquisitions are released within ``unprocessed'' packages, and
  the ``minimally preprocessed'' are corrected for artifacts, brain-extracted
  and spatially normalized, along with some results of the standard processing
  pipeline of \emph{FreeSurfer}.

@@

Conversely, the susceptibility-derived distortions only happen along the \gls*{pe}
  axis of the image.
Therefore, a \gls*{swindex} can be computed as the one-to-one distance between corresponding
  vertices of surfaces, weighted by their respective Voronoi area $a_i$:

  \begin{equation}
  sWI = \frac{1}{\sum_i a_i} \sum\limits_i^{N_V} a_i\,\|
  \vec{v}_i - \hat{\vec{v}}_i \|
  \label{eq:swindex}
  \end{equation}
%
  where $\vec{v}_i \subset \gammaset$ are the locations of the total $N_V$ vertices, and $\hat{\vec{v}}_i$ is the location
  recovered corresponding to the vertex $\vec{v}_i$.
